---
title: "Clustering Analysis Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clustering Analysis Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(GeneClustR)
library(ggplot2)
library(plotly)
library(cluster)
library(factoextra)
```

This vignette shows how to determine the optimal number of clusters and how to apply different clustering algorithms on pre-processed gene expression data.

First of all, let's build a 'simulated mtrix' in order to show the functionality of this block.

```{r}
set.seed(123)
mat <- matrix(rnorm(300), nrow = 30, ncol = 10)
rownames(mat) <- paste0("Gene", 1:30)
colnames(mat) <- paste0("Sample", 1:10)

# Preprocessing
mat_prep <- preprocess(mat, normalize = TRUE, method = "log")

# Rimuovi righe con valori non finiti (Inf, -Inf, NA, NaN)
mat_prep <- mat_prep[apply(mat_prep, 1, function(x) all(is.finite(x))), ]
```

Determination of the number of clusters:

Elbow Method, the default one, require a matrix/dataframe in input and the desired number of clusters:

```{r}
cluster_number(mat_prep, method = "elbow", max_clusters = 10)

```

Gap Statistics method:

```{r}
cluster_number(mat_prep, method = "gapstats", max_clusters = 10)

```

Once you have runned the (optional) function, used to have an indication on the 'optimal' cluster number, is time to choose a clustering method used to obtain a list containing useful summary and information and also plots

k-means:

```{r}
kmeans_res <- clustering(mat_prep, method = "kmeans", centers = 3)
kmeans_res$plot

```

k-medoids

```{r}
kmedoids_res <- clustering(mat_prep, method = "kmedoids", centers = 3, metric = "euclidean")
kmedoids_res$plot

```

hierarchical clustering

```{r}
hclust_res <- clustering(mat_prep, method = "hierarchical", centers = 3)
hclust_res$plot  # PCA view
hclust_res$dendrogram  # Dendrogramma

```

DBSCAN

```{r}
# eps e minpts vanno trovati tramite sperimentazione (es. kNNdistplot)
dbscan_res <- clustering(mat_prep, method = "DBSCAN", eps = 1, minpts = 2)
dbscan_res$plot

```

